---
title: "Final Project"
output: 
   html_document:
      theme:
         journal
runtime: shiny
date: '2022-09-30'
---

<font size="4">\

```{=html}
<style>
body {background-color: #ffe4e1}
</style>
```
In this project, I used "Students Performance in Exams" data set in order to make an interesting visualization to better understand the data.

```{r chunk1, echo = FALSE, warning = FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
data_1<-read.csv("D:\\MA\\final_project\\StudentsPerformance.csv")

```

#### This is my raw data

```{r chunk2, echo = FALSE, warning = FALSE, message=FALSE}
head(data_1)
```

#### First I am going to organize the data

#### Let's rename the factorial levels of the data:

```{r chunck3, warning = FALSE, message=FALSE}
data_1$parental.level.of.education<-as.factor(data_1$parental.level.of.education)
levels(data_1$parental.level.of.education)
data_1$parental.level.of.education <- recode_factor(data_1$parental.level.of.education, "some high school" = "high school", "some college" = "college")  

```

This is an interactive table, which can help us in understanding and "feeling" the data 

We can filter the data to see the data in different conditions

```{r chunk4, echo = FALSE, warning = FALSE, message=FALSE}
library(shiny)
library(shinydashboard)
library(DT)
library(ggplot2)
library(yaml)
library(knitr)
library(rmarkdown)


ui<-fluidPage(
  titlePanel("DataTable"),
  fluidRow(
    column(4,
        selectInput("gender",
                    "Gender",
                    c("All",
                      unique(as.character(data_1$gender))))
    ),
    column(4,
        selectInput("lunch",
                    "Lunch",
                    c("All",
                      unique(as.character(data_1$lunch))))
    ),
    column(4,
        selectInput("parental.level.of.education",
                    "Parental level of education",
                    c("All",
                      unique(as.character(data_1$parental.level.of.education))))
    )
  ),

  DT::dataTableOutput("table")
)

server<- function(input, output) {

  # Filter data based on selections
  output$table <- DT::renderDataTable(DT::datatable({
    data <-data_1
    if (input$gender != "All") {
      data <- data[data$gender == input$gender,]
    }
    if (input$lunch != "All") {
      data <- data[data$lunch == input$lunch,]
    }
    if (input$parental.level.of.education != "All") {
      data <- data[data$parental.level.of.education == input$parental.level.of.education,]
    }
    data
  }))

}

shinyApp(ui, server)
```

After playing with the data, we can see that pupils who completed preparation and that their parents are educated are more likely to get higher scores

Next, we let's show the data in bar chart:

```{r chunk5, echo = FALSE, warning = FALSE, message=FALSE}

library(shiny)
library(dplyr)

server<- function(input, output, session) {

data1<- reactive({
  req(input$select_preparation)
  df <- data_1 %>% filter(test.preparation.course %in% input$select_preparation) %>% group_by(gender) %>% summarize(mean(math.score,na.rm=TRUE), mean(writing.score,na.rm=TRUE), mean(reading.score,na.rm=TRUE))
df <- as.matrix(df[,-1])
})

#update SelectInput
observe({
  updateSelectInput(session, "select_preparation", choices = data_1$test.preparation.course)
})

#Plot
output$plot<- renderPlot({
barplot(data1() , beside = T , legend.text = T,col = c("blue" , "skyblue"), xlab = "subject of study", ylab = "mean scores",names.arg = c("Math", "Writing", "Reading"))})
}

ui<- basicPage(
h1("Shiny with interactive bar chart")  ,
selectInput(inputId = "select_preparation",
label="Choose preparation", 
"peraparation"),
plotOutput("plot")
  )

shinyApp(ui, server)
```

</font>


